---
title: "Figure 2C 3"
author: "Emma Chapman-Banks"
date: "2025-01-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com"))
```

# ***Part 1: Cross-Referencing Policy Documents and Abstracts (Same Country)*** 

Load libraries
```{r}
library("tidyverse")
library("sf") 
library("countrycode") 
library("ggrepel") 
library("gt")
library("dplyr")
library("tidyr")
library(rnaturalearth)
```

Read the data
```{r}
overton_data <- read.csv("~/Google Drive/My Drive/HITAP/IDM Literature/analysis/data/data/overton_results_expanded.csv")
idm_data <- read.csv("~/Google Drive/My Drive/HITAP/IDM Literature/analysis/data/data/clean_idm_lit_with_locations.csv")

overton_data <- overton_data[, !grepl("^X", names(overton_data))]
idm_data <- idm_data[, !grepl("^X", names(idm_data))]

```

World Data
```{r}
# Get world data 
world <- ne_countries(scale = "small", returnclass = "sf")
world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$iso_a3_eh,
                       world$iso_a3)

world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$sov_a3,
                       world$iso_a3)

# Fix iso_a3 for France
world <- world %>%
  mutate(iso_a3 = ifelse(name == "France", "FRA", iso_a3))

# Define ISO 3 codes
iso3_codes <- c("AND", "BHR", "COK", "COM", "CPV", "CUW", "FRO", "FSM", "GLP", "GUF", 
                "GUM", "HKG", "IMN", "JEY", "LCA", "LIE", "MAC", "MDV", "MTQ", "MUS", 
                "MYT", "NFK", "PYF", "REU", "SGP", "STP", "SYC", "unmapped", "VIR", 
                "WSM")

# Create a mapping for UN regions
un_region_mapping <- list(
  "Europe" = c("AND", "FRO", "IMN", "JEY", "LIE"),
  "Asia" = c("BHR", "HKG", "MAC", "MDV", "SGP"),
  "Oceania" = c("COK", "FSM", "GUM", "NFK", "PYF", "WSM"),
  "Africa" = c("COM", "CPV", "MYT", "REU", "STP", "SYC", "MUS"),
  "Americas" = c("CUW", "GLP", "GUF", "MTQ", "LCA", "VIR"),
  "Antarctica" = character(0)  # No codes for Antarctica in this list
)

# Create a mapping for Income Group based on World Bank 2019 classification
inc_group_mapping <- list(
  "2. High income: nonOECD" = c("AND", "BHR", "CUW", "FRO", "GLP", "GUM", "HKG", "IMN", "JEY", 
                    "LIE", "MAC", "MTQ", "MUS", "REU", "SGP"),
  "3. Upper middle income" = c("CPV", "MDV", "STP", "SYC", "LCA"),
  "4. Lower middle income" = c("COM", "FSM"),
  "5. Low income" = character(0),  # No codes in this category
  "Unmapped" = c("unmapped")
)

# Function to map values
map_values <- function(code, mapping) {
  match <- unlist(lapply(names(mapping), function(region) if (code %in% mapping[[region]]) region else NULL))
  if (length(match) == 0) NA else match
}

# Create dataset
df <- data.frame(
  iso_a3 = iso3_codes,
  region_un = sapply(iso3_codes, map_values, mapping = un_region_mapping),
  income_grp = sapply(iso3_codes, map_values, mapping = inc_group_mapping),
  stringsAsFactors = FALSE
)

# Define ISO 3 codes and country names
iso3_names <- c(
  "AND" = "Andorra", "BHR" = "Bahrain", "COK" = "Cook Islands", "COM" = "Comoros",
  "CPV" = "Cape Verde", "CUW" = "Curaçao", "FRO" = "Faroe Islands", "FSM" = "Micronesia",
  "GLP" = "Guadeloupe", "GUF" = "French Guiana", "GUM" = "Guam", "HKG" = "Hong Kong",
  "IMN" = "Isle of Man", "JEY" = "Jersey", "LCA" = "Saint Lucia", "LIE" = "Liechtenstein",
  "MAC" = "Macau", "MDV" = "Maldives", "MTQ" = "Martinique", "MUS" = "Mauritius",
  "MYT" = "Mayotte", "NFK" = "Norfolk Island", "PYF" = "French Polynesia", "REU" = "Réunion",
  "SGP" = "Singapore", "STP" = "São Tomé and Príncipe", "SYC" = "Seychelles",
  "VIR" = "U.S. Virgin Islands", "WSM" = "Samoa", "unmapped" = NA
)

# Add country names to df
df <- df %>%
  mutate(admin = iso3_names[iso_a3])  # Match country names based on ISO 3 codes

world <- bind_rows(world, df)
```

Before creating this R markdown, I ran the analysis by joining the datasets together then cross-referencing the dois and country location. This returned zero results. At first, I was puzzled, but then I realised it was because the countries might be spelt differently or be abbreviated. Therefore, before joining the datasets, I used the Iso3 country codes and standardised the country names across the data sets 

Standardise IDM Data. This code is the same code used in the World Mapping Markdown.

When trying to standarise the country codes for Overton data, I also realised that the new column it creates returns a value of NA if the country code is already written. Thus, I identified which country codes were already listed in the dataset and had to include the manual country mapping and ensure it doesn't pass as NA when doing country code mapping. 

```{r}
# Define manual mappings for unmatched values
manual_country_mapping <- c(
  "IGO" = "IGO",  
  "EU" = "EU"     
)

# Standardise Overton Data and include manual mapping
overton_data_clean <- overton_data %>%
  mutate(
    country_iso3 = if_else(
      country %in% names(manual_country_mapping),
      manual_country_mapping[country],
      countrycode::countrycode(
        sourcevar = country,
        origin = "country.name",
        destination = "iso3c",
        warn = FALSE 
      )
    )
  )

```

Now, cross-reference the two datasets.
```{r}
# Combine the datasets based on DOI and use the ISO3 country codes for comparison
combined_data <- overton_data_clean %>%
  inner_join(idm_data, by = c("source_doi" = "doi"))

combined_data <- combined_data[, !grepl("^X", names(combined_data))]

# Filter where the policy document's country matches the IDM abstract's location
matching_countries <- combined_data %>%
  filter(country_iso3 == Iso3) %>%
  filter(Status != "Unmapped")

# Count the number of matches
number_of_matches <- nrow(matching_countries)

# Print the result
print(number_of_matches)

# Summarize by Iso3 (count the number of abstracts per country)
matching_countries_map <- matching_countries %>%
  group_by(Iso3) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(matching_countries_map, by = c("iso_a3" = "Iso3"))

# Define Lancet color palette for continuous fill
lancet_colors <- pal_lancet()(2)  # Get 9 colors from Lancet palette

# Plot the world map with Lancet colors
ggplot(world_with_data) +
  geom_sf(aes(fill = count)) +
  scale_fill_gradientn(
    colors = lancet_colors, 
    na.value = "grey90", 
    name = "Abstract Count"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_from_their_own_countries.jpg")
```

# ***Part 2: Cross-Referencing Policy Documents and Abstracts (Different Country)*** 

Now, let's answer this question: how many policy documents cite IDM literature from a country different from their own?
```{r}
# Filter where the policy document cites IDM literature from a different country
global_research_citations <- combined_data %>%
  filter(country_iso3 != Iso3)  %>%
  filter(Status != "Unmapped")

# Count the number of matches
num_global_citations <- nrow(global_research_citations)

# Print the result
print(num_global_citations)

# Summarize by Iso3 (count the number of abstracts per country)
global_research_citations_map <- global_research_citations %>%
  group_by(Iso3) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(global_research_citations_map, by = c("iso_a3" = "Iso3"))

# Define Lancet color palette for continuous fill
lancet_colors <- pal_lancet()(2)  # Extract 9 colors from the Lancet palette

# Plot the world map with Lancet colors
ggplot(world_with_data) +
  geom_sf(aes(fill = count)) +
  scale_fill_gradientn(
    colors = lancet_colors, 
    na.value = "grey90", 
    name = "Abstract Count"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")


ggsave("plots/citing_from_other_countries.jpg")
```

# ***Part 3: Deviance between self-citing & other countries*** 

Now, let's answer this question: what's the difference in citing national work versus citing foreign work?
```{r}
# Filter where the policy document cites IDM literature from a different country
names(global_research_citations_map) <- c("Iso3", "count_foreign_citations")
names(matching_countries_map) <- c("Iso3", "count_national_citations")

combined_data_deviance <- merge(matching_countries_map, global_research_citations_map, by = "Iso3", all = TRUE)

combined_data_deviance$diff <- combined_data_deviance$count_national_citations - combined_data_deviance$count_foreign_citations

combined_data_deviance$ratio <- combined_data_deviance$count_national_citations/ combined_data_deviance$count_foreign_citations

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(combined_data_deviance, by = c("iso_a3" = "Iso3"))

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = diff)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Citations from foreign IDM vs. national IDM literature") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_deviance.jpg")

# Define Lancet color palette for continuous fill
lancet_colors <- pal_lancet()(2)  # Extract 9 colors from the Lancet palette

world_with_data <- world_with_data %>%
  filter(name != "Antarctica")

# Plot the world map with Lancet colors
ggplot(world_with_data) +
  geom_sf(aes(fill = ratio)) +
  scale_fill_gradientn(
    colors = lancet_colors, 
    na.value = "grey90", 
    name = "Ratio of Citations from foreign IDM to domestic IDM literature"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/Fig2B_citing_ratio.jpg")
```

```{r}

# Convert the date column to Date format (optional but recommended)
overton_data_clean$published_on <- as.Date(overton_data_clean$published_on)

# Extract the year and overwrite the column
overton_data_clean$published_on <- format(overton_data_clean$published_on, "%Y")

pc <- overton_data_clean |>
  group_by(published_on, country_iso3) |>
  summarize(n = n()) |>
  rename(n_cit = n)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(pc, by = c("iso_a3" = "country_iso3"))

world_with_data <- world_with_data %>%
  dplyr::filter(!is.na(published_on))

# Plot the bar chart with Lancet colors
ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = economy)) +
  geom_bar(stat = "identity") +
  scale_fill_lancet(name = "Income Group") +  # Apply Lancet colors for discrete fill
  labs(
    x = "Year", 
    y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
  ) + 
  theme_minimal()

ggsave("plots/by_economy.png",
       width = 15)

# Plot the bar chart with Lancet colors
ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
  geom_bar(stat = "identity") +
  scale_fill_lancet(name = "Income Group") +  # Apply Lancet colors for discrete fill
  labs(
    x = "Year", 
    y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
  ) + 
  theme_minimal()



ggsave("plots/by_income_group.png",
       width = 15)


```


```{r}
1 - length(unique(overton_data_clean$country))/195
pc <- overton_data_clean |>
  group_by(published_on, country_iso3, type) |>
  summarize(n = n()) |>
  rename(n_cit = n)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))

pc_igo <- pc %>% 
  filter(type == "igo")

pc_igo$income_grp <- "Not country-specific"

names(pc_igo) <- c("published_on", "iso_a3", "type.y","n_cit","income_grp")

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(pc, by = c("iso_a3" = "country_iso3"))

world_with_data <- world_with_data[,c("published_on", "iso_a3", "type.y","n_cit","income_grp")]
world_with_data <- st_drop_geometry(world_with_data)

world_with_data <- bind_rows(world_with_data, pc_igo)

world_with_data <- world_with_data %>%
  dplyr::filter(!is.na(published_on))

world_with_data$type.y <- factor(world_with_data$type.y, levels = c("government", "think tank", "igo","other"))

# Plot the bar chart with Lancet colors
ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
  geom_bar(stat = "identity") +
  scale_fill_lancet(name = "Income Group") +  # Apply Lancet colors for discrete fill
  labs(
    x = "Year", 
    y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
  ) + 
  facet_grid(rows = vars(type.y)) +  # Stack facets vertically based on type.y
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis text by 45 degrees
  )


ggsave("plots/by_income_group_type.png",
       width = 14)


library(ggplot2)
library(ggsci)  # For Lancet colors
library(viridis)  # Optional alternative color scheme

world_with_data$type.y <- gsub("government", "Government", world_with_data$type.y)
world_with_data$type.y <- gsub("think tank", "Think Tank", world_with_data$type.y)
world_with_data$type.y <- gsub("other", "Other", world_with_data$type.y)
world_with_data$type.y <- gsub("igo", "IGO", world_with_data$type.y)

ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
  geom_bar(stat = "identity") +
  
  scale_fill_lancet(name = "Income Group") +  # Apply Lancet colors for discrete fill
  
  labs(
    x = "Year", 
    y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
  ) + 
  
  facet_grid(rows = vars(type.y)) +  # Stack facets vertically based on type.y
  
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),  # Add box around graph
    panel.grid.major = element_line(color = "grey90", linewidth = 0.25),  # Lighter major grid
    panel.grid.minor = element_line(color = "grey95", linewidth = 0.25),  # Lighter minor grid
    strip.text = element_text(size = 14, face = "bold"),  # Make facet labels slightly larger & bold
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12),  # Rotate x-axis text for readability
    axis.text.y = element_text(size = 12),  # Adjust y-axis text size
    axis.title = element_text(size = 14),  # Increase axis label size
    legend.position = "bottom",  # Keep legend at bottom for better readability
    legend.title = element_text(size = 13, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 13)
  )
  
  
ggsave("plots/by_income_group_type2.png",
       width = 14, height = 7)

country_df <- world_with_data |>
  group_by(iso_a3,type.y, income_grp) |>
  summarize(n = sum(n_cit)) |>
  rename(n_cit = n)

write.csv(country_df, "results/country_df_citations.csv",row.names = FALSE)
```

