---
title: "Code for Figure 1: Timeline"
author: "Emma Chapman-Banks"
date: "2025-01-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ***Part 1: Merging Figures 1 and 4***

Load necessary packages
```{r}
library(ggplot2)
library(dplyr)
library(janitor)
library(ggsci)
```

- read, clean, and save publication data.
```{r}
# Get Publication data
df <- read.csv("~/Google Drive/My Drive/HITAP/IDM Literature/analysis/data/data/490k_141224_1234_dedup.csv") #Or change file path to wherever files are located
df <- clean_names(df)

# Remove all excluded records
df <- df |>
  filter(decision == "Include")

# Calc publication records by year 
df_annual <- df |>
  group_by(year) |>
  summarize(n = n()) |>
  rename(n_pub = n)

# Remove records prior to 1900 or after 2025 
df_annual <- df_annual |>
  filter(year < 2025) |>
  filter(year > 1900)

df_annual_save <- df_annual
```

- load in data for policy citations
```{r}

pc <- read.csv("~/Google Drive/My Drive/HITAP/IDM Literature/analysis/data/data/overton_results_expanded.csv") 
pc <- clean_names(pc)

# Transform date of policy documents to year get year 
pc$published_on <- as.Date(pc$published_on)
pc$year <- format(as.Date(pc$published_on, format="%Y-%m/%d"),"%Y")

# Calc policy citations by year 
pc_annual <- pc |>
  group_by(year) |>
  summarize(n = n()) |>
  rename(n_cit = n)

# Calc policy citations by org type and subtype
pc_subtype <- pc |>
                group_by(type, subtype) |>
                summarize(n = n()) |>
                rename(n_cit = n)

write.csv(pc_subtype, 
          "plots/ec_aggregation_by_sub_type.csv", 
          row.names = FALSE)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))
```


- pc_annual starts from 1989 and df_annual starts from 1948. This can create issues when merging so we must align the data ranges to ensure RStudio can read this correctly. 
```{r}
# Merge datasets, keeping all years from df_annual
combined_df <- merge(df_annual, pc_annual, by = "year", all.x = TRUE)

# Replace missing policy citation values (n_cit) with 0 (as pc_annual doesn't begin until 1989)
combined_df$n_cit[is.na(combined_df$n_cit)] <- 0

# Ensure n_pub is numeric
combined_df$n_pub <- as.numeric(combined_df$n_pub)
```

# ***Part 2: Merging all three plots together***

Now, we will "superimpose" the publication count/policy citations with the outbreak timeline. 

But first, lets load the outbreak timeline data (Paula's code)

```{r}
# Get Outbreaks timeline data 
e_timeline <- readxl::read_excel("data/Epidemics Timeline.xlsx", 
                                 sheet = 2)
names(e_timeline)[1] <- c("year")
e_timeline <- clean_names(e_timeline)

# Reshape data to long format for ggplot2 
data_long <- tidyr::pivot_longer(e_timeline, 
                                 -year, 
                                 names_to = "pathogen", 
                                 values_to = "presence")
data_long <- data_long |> filter(presence > 0)
data_long <- data_long |> filter(year >= min(df_annual$year))

# Renaming pathogens. This is done to help easier visualisation on the plot 
rename_mapping <- c(
  "zika" = "Zika",
  "swine_flu_h1n1_pandemic" = "H1N1 Swine Flu",
  "severe_acute_respiratory_syndrome_sars_coronavirus" = "SARS",
  "h1n1_russian_flu" = "H1N1 Russian Flu",
  "m_pox" = "MPox",
  "hong_kong_flu_pandemic_h3n2" = "H3N2 Hong Kong Flu",
  "asian_flu_pandemic_h2n2" = "H2N2 Asian Flu",
  "mers" = "MERS",
  "covid_19" = "COVID-19",
  "ebola" = "Ebola",
  "uptick_in_polio" = "Polio",
  "hiv_aids_pandemic" = "HIV/AIDS"
)

data_long$pathogen <- dplyr::recode(data_long$pathogen, !!!rename_mapping)

# Reorder factor levels by frequency 
data_long$pathogen <- factor(data_long$pathogen, 
                             levels = names(sort(table(data_long$pathogen), 
                                                 decreasing = TRUE)))

# Order dataframe by year
data_long <- data_long[order(data_long$year), ]
```

We will now add the outbreak plot onto the publication/policy plot.

This is the old plot (no disease y-axis with circle dot points)
```{r}
#Because I had issues with creating the legend. I needed to convert the legend labels into a factor (for pathogens) then precompute the breaks 
data_long$pathogen <- factor(data_long$pathogen)
pathogen_breaks <- c("Publication Count", "Policy Citations", levels(data_long$pathogen))

```


Updated Plot (Jan 24th): adding labels to the bars. 

Issue here: the horizontal bars are made up of multiple individual rectangles. You cannot see it in the graph because it has been adjusted so that each rectangle is the same height and width. Therefore, what I have done is pass a code first the calculates the central y position and average x position of each "rectangle". So you must pass this code before plotting it. 
```{r}
# Making sure data is numeric 
data_long <- data_long %>%
  mutate(outbreak_y = as.numeric(as.factor(pathogen))) 

first_appearance <- data_long %>%
  group_by(pathogen) %>%
  summarise(first_year = min(year)) %>%
  arrange(first_year)

# Re-level the pathogen factor in data_long and then create outbreak_y
data_long <- data_long %>%
  mutate(pathogen = factor(pathogen, levels = first_appearance$pathogen),
         outbreak_y = as.numeric(pathogen))

# Calculate the central position for the labels
centered_data <- data_long %>%
  group_by(pathogen) %>%
  summarise(
    center_y = mean(outbreak_y * 300),  # This is the center of the horizontal bar
    center_x = mean(year)              # This is the center x position (average year)
  )

centered_data$center_x <- ifelse(centered_data$pathogen == "Ebola", 2016.4, centered_data$center_x)

ebola_add <- data.frame(pathogen = "Ebola",
                        center_y = 1500,
                        center_x = 2000)
centered_data <- rbind(centered_data, ebola_add)
```

```{r}



ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1) +
  geom_tile(
    data = data_long,
    aes(x = year, y = outbreak_y * 300),  # No fill mapping, just a fixed color
    fill = "lightgrey", 
    width = 1, height = 200,
    alpha = 0.3
  ) +
  geom_text(
    data = centered_data,
    aes(
      x = center_x,          
      y = center_y,          
      label = pathogen
    ),
    vjust = 0.5,             
    hjust = 0.5,             
    size = 3,
    inherit.aes = FALSE,
    color = "black"
  ) +
  geom_vline(xintercept = c(1989, 2018, 2021), linetype = "dashed", color = "darkgrey", size = 1) +
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  scale_x_continuous(
    limits = c(1960, NA),  # Set lower limit to 1960, upper limit remains automatic
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  
  # Apply Lancet color palette from ggsci for lines
  scale_color_lancet(name = " ") +  
  
  labs(x = "Year", color = " ") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.25),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.25)
  )

ggsave("plots/timeline.png", width = 10)
```

```{r}
library(ggplot2)
library(ggsci)  # Lancet color palette
library(viridis)  # Alternative color palettes

ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1.5) +  # Thicker lines
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1.5) +  
  
  geom_tile(
    data = data_long,
    aes(x = year, y = outbreak_y * 300),  
    fill = "lightblue",  # Slightly more distinct outbreak markers
    width = 1, height = 200,
    alpha = 0.3
  ) +
  
  geom_text(
    data = centered_data,
    aes(
      x = center_x,          
      y = center_y,          
      label = pathogen
    ),
    vjust = 0.5,             
    hjust = 0.5,             
    size = 4,  # Slightly larger text
    inherit.aes = FALSE,
    color = "black"
  ) +
  
  geom_vline(xintercept = c(1989, 2018, 2021), linetype = "dashed", color = "darkgrey", size = 1) +  # More distinct color
  
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  
  scale_x_continuous(
    limits = c(1960, NA),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  
  scale_color_lancet(name = " ") +  # Lancet colors for clarity
  
  labs(x = "Year", color = " ") +
  
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "black", size = 14),
    axis.text.y.right = element_text(color = "black", size = 12),
    axis.title.x = element_text(size = 14),        # Increased X title
    axis.text.x = element_text(size = 12),
    axis.title.y.left = element_text(color = "black", size = 14),
    axis.text.y.left = element_text(color = "black", size = 12),

    panel.grid.major = element_line(color = "grey90", linewidth = 0.25),  # Lighter grid
    panel.grid.minor = element_line(color = "grey95", linewidth = 0.25),  # Even lighter minor grid
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),  # Box around graph
    plot.background = element_rect(fill = "white", color = NA),
    legend.title = element_text(size = 13, face = "bold"),  # Bold legend title
    legend.text = element_text(size = 13)
  )
  
ggsave("plots/timeline2.png", width = 10)

```


