---
title: 'R Markdown (UPDATED: World Mapping)'
author: "Emma Chapman-Banks"
date: "2025-01-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ***Part 1: Load packages/set up world map***

Load necessary packages
```{r}
library("tidyverse") 
library("sf") 
library("rnaturalearth") 
library("countrycode")
library("ggrepel") 
library("stringr")
```
World Data
```{r}
# Get world data 
world <- ne_countries(scale = "small", returnclass = "sf")
world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$iso_a3_eh,
                       world$iso_a3)

world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$sov_a3,
                       world$iso_a3)

# Fix iso_a3 for France
world <- world %>%
  mutate(iso_a3 = ifelse(name == "France", "FRA", iso_a3))
```
# ***Part 2: Identify Unmapped Locations***

```{r}
# Load new dataset with geolocations. 
abstracts_data_clean <- read.csv("data/Combined_Location_With_Metadata_All.csv", stringsAsFactors = FALSE)

abstracts_data_clean <- abstracts_data_clean[, !grepl("^X", names(abstracts_data_clean))]

# Process `location` column
abstracts_data_clean <- abstracts_data_clean %>%
  filter(!is.na(location) & location != "") %>%
  separate_rows(location, sep = ";\\s*") %>%  # Split multiple locations into rows
  mutate(
    Iso3 = countrycode::countrycode(
      sourcevar = location,
      origin = "country.name",
      destination = "iso3c",
      warn = FALSE  # Suppress warnings for unmatched entries
    ),
    Iso3 = ifelse(is.na(Iso3), "unmapped", Iso3)  # Mark unmatched values as "unmapped"
  ) %>%
  mutate(Status = ifelse(Iso3 == "unmapped", "Unmapped", "Mapped"))  

# Look at unmatched values
unmatched <- abstracts_data_clean %>%
  filter(Status == "Unmapped") %>%
  distinct(location) %>%
  arrange(location)

# Print unmatched values for verification
#print(unmatched)

# Look at matched values
matched <- abstracts_data_clean %>%
  filter(Status == "Mapped") %>%
  distinct(location, Iso3) %>%
  arrange(Iso3)

# Print matched values 
#print(matched)
```
# ***Part 3: Clean and re-map unmapped entries***

After identifying the list of unmapped entries, I did a quick scan through to understand how I could process the values. I decided to group all the values by country as some of the locations cited were cities/regions/places/hospitals that the Iso3 code would not pick up. 

View unmatched entries in the US 
```{r}
# Filter for US-related unmatched entries
unmatched_us <- unmatched %>%
  filter(str_detect(location, regex("U\\.S\\.|US|USA|United States|North America", ignore_case = TRUE)))

# Preview the unmatched US entries
#print(unmatched_us)
```
View unmatched entries in the UK 
```{r}
# Filter for UK-related unmatched entries
unmatched_uk <- unmatched %>%
  filter(str_detect(location, regex("U\\.K\\.|UK|United Kingdom|England|Scotland|Ireland|Wales|Welsh|British Isles|Tristan da Cunha", ignore_case = TRUE)))

# Preview the unmatched UK entries
#print(unmatched_uk)
```
Identify any rows that mention US states and other cities/counties (as some don't explictly say US/USA/U.S)
```{r}
# List and define US States
us_states <- paste(
  c(
    "Alabama", "Alaska", "Arizona", "Arkansas", "California", "Colorado", "Connecticut", "Delaware", "Florida", "Georgia", "Hawaii", "Idaho", "Illinois", "Indiana", "Iowa", "Kansas", "Kentucky", "Louisiana", "Maine", "Maryland", "Massachusetts", "Michigan", "Minnesota", "Mississippi","Missouri", "Montana", "Nebraska", "Nevada", "New Hampshire", "New Jersey", "New Mexico","New York", "North Carolina", "North Dakota", "Ohio", "Oklahoma", "Oregon", "Pennsylvania","Rhode Island", "South Carolina", "South Dakota", "Tennessee", "Texas", "Utah", "Vermont","Virginia", "Washington", "West Virginia", "Wisconsin", "Wyoming", "Los Angeles" , "Miami", "New Haven", "Northeastern University", "Philadelphia", "Atlanta", "Barnes-Jewish Hospital", "Detroit Medical Center", "Johns Hopkins HIV Clinical Cohort", "M. D. Anderson Cancer Center", "Mayo Clinic Hospitals", "Mayo Clinic enterprise hospitals", "Mount Sinai Health System", "O'ahu, Hawai'i", "Seattle", "Vanderbilt University", "Mass General Brigham", "Maui, Hawai'i", "Greenville Health System Hernia Center", "M Health Fairview", "Navajo Nation", "Saint Martin"),
  collapse = "|"
)

# Filter for rows mentioning any US state
unmatched_usstates <- unmatched %>%
  filter(str_detect(location, regex(us_states, ignore_case = TRUE)))

# Preview the unmatched state-related entries
#print(unmatched_usstates)
```

Now, let's do China (and chinese cities + beijing)
```{r}
# Filter for China unmatched entries
unmatched_china <- unmatched %>%
  filter(str_detect(location, regex("China|Chinese cities|Chinese Cities|Beijing|Chinese provinces|Guangdong Province|Hangzhou, Zhejiang Province|Hubei province|Zhejiang Province|Kunming|Yunnan Province|Hunan|Wuhan|Xinjiang region|Yangtze River Delta|Guangdong|Guangxi|Hubei|Fuyang|Qinling-Huaihe Line|Anhui|Chinese mainland|The 904th Hospital of the Joint Logistics Support Force of the PLA" , ignore_case = TRUE)))

# Preview the unmatched China entries
#print(unmatched_china)
```

View Candian cities/provinces
```{r}
# Filter for Canadian unmatched entries
unmatched_canada <- unmatched %>%
  filter(str_detect(location, regex("Canadian|Quebec|Ontario|Alberta|British Columbia|Winnipeg, Manitoba|Manitoba|Nova Scotia|Saskatchewan|Newfoundland and Labrador", ignore_case = TRUE)))

# Preview the unmatched Canada entries
#print(unmatched_canada)
```

Group Japan-associated areas 
```{r}
# Filter for Japanese unmatched entries
unmatched_japan <- unmatched %>%
  filter(str_detect(location, regex("Kanagawa|Osaka|Tokyo", ignore_case = TRUE)))

# Preview the unmatched Japanese entries
#print(unmatched_japan)
```

Group Australia-associated areas 
```{r}
# Filter for Australia unmatched entries
unmatched_australia <- unmatched %>%
  filter(str_detect(location, regex("Melbourne, Victoria|Queensland|Tasmania|Victoria" , ignore_case = TRUE)))

# Preview the unmatched Australian entries
#print(unmatched_australia)
```

Identiy Yemen-associated areas 
```{r}
# Filter for Yemen unmatched entries
unmatched_yemen <- unmatched %>%
  filter(str_detect(location, regex("Yemen|Sana'a, Yemen|Yemen Arab Republic" , ignore_case = TRUE)))

# Preview the unmatched Yemen entries
#print(unmatched_yemen)
```

Identiy Brazil-associated areas 
```{r}
# Filter for Brazil unmatched entries
unmatched_brazil <- unmatched %>%
  filter(str_detect(location, regex("Amazon|Amazon Basin|Amazonian", ignore_case = TRUE)))

# Preview the unmatched Brazil entries
#print(unmatched_brazil)
```

View Indian-associated areas
```{r}
# Filter for Indian unmatched entries
unmatched_india <- unmatched %>%
  filter(str_detect(location, regex("Andhra Pradesh|Benga|Bengal Delta|Delhi|Gujarat|Karnataka|Kerala|Maharashtra|Tamil Nadu", ignore_case = TRUE)))

# Preview the unmatched Indian entries
#print(unmatched_india)
```

View Italian-associated areas
```{r}
# Filter for Italian unmatched entries
unmatched_italy <- unmatched %>%
  filter(str_detect(location, regex("Apulia|Basilicata|Calabria|Campania|Emilia Romagna|Lazio|Lombardia|Lombardy|Marche|Molise|Piedmont|Sicily|Abruzzo|San Raffaele Hospital|Veneto", ignore_case = TRUE)))

# Preview the unmatched Italian entries
#print(unmatched_italy)
```

View Thai-associated areas
```{r}
# Filter for Thai unmatched entries
unmatched_thai <- unmatched %>%
  filter(str_detect(location, regex("Bangkok|Phramongkutklao Hospital|rural Thai settings", ignore_case = TRUE)))

# Preview the unmatched Thai entries
#print(unmatched_thai)
```

View Indonesia-associated areas
```{r}
# Filter for Indonesia unmatched entries
unmatched_indonesia <- unmatched %>%
  filter(str_detect(location, regex("Banjarmasin, South Kalimantan|Papua, Indonesia|West Papua, Indonesia", ignore_case = TRUE)))

# Preview the unmatched Indonesia entries
#print(unmatched_indonesia)
```

View Spanish-associated areas
```{r}
# Filter for Spanish unmatched entries
unmatched_spain <- unmatched %>%
  filter(str_detect(location, regex("Basque Country|Catalonia|Lleida, Catalonia", ignore_case = TRUE)))

# Preview the unmatched Spanish entries
#print(unmatched_spain)
```

View French-associated areas
```{r}
# Filter for French unmatched entries
unmatched_france <- unmatched %>%
  filter(str_detect(location, regex("French West Indies|Institut du Cancer de Montpellier|Reunion Island, France|French alpine villages", ignore_case = TRUE)))

# Preview the unmatched French entries
#print(unmatched_france)
```

View Korea-associated areas
```{r}
# Filter for Korea unmatched entries
unmatched_korea <- unmatched %>%
  filter(str_detect(location, regex("Daegu, Republic of Korea|Daegu-Kyungpook Province, South Korea", ignore_case = TRUE)))

# Preview the unmatched Korea entries
#print(unmatched_korea)
```

View Tanzania-associated areas
```{r}
# Filter for Tanzania unmatched entries
unmatched_tanzania <- unmatched %>%
  filter(str_detect(location, regex("Rufiji, Tanzania|Pemba, Zanzibar|Pemba Island, Zanzibar|Zanzibar", ignore_case = TRUE)))

# Preview the unmatched Tanzania entries
#print(unmatched_tanzania)
```

View Switzerland-associated areas
```{r}
# Filter for Switzerland unmatched entries
unmatched_switzerland <- unmatched %>%
  filter(str_detect(location, regex("Lake Geneva|University Hospital of Zurich", ignore_case = TRUE)))

# Preview the unmatched Switzerland entries
#print(unmatched_switzerland)
```

View Micronesia-associated areas
```{r}
# Filter for Micronesia unmatched entries
unmatched_micronesia <- unmatched %>%
  filter(str_detect(location, regex("Fais, Micronesia|Yap, Micronesia|Yap Island|Yap Island, Oceania|Yap Main Islands, Micronesia", ignore_case = TRUE)))

# Preview the unmatched Micronesian entries
#print(unmatched_micronesia)
```

View Saudi Arabia-associated areas
```{r}
# Filter for Saudi unmatched entries
unmatched_saudi <- unmatched %>%
  filter(str_detect(location, regex("Gazan Province, Saudi Arabia|KSA", ignore_case = TRUE)))

# Preview the unmatched Saudi entries
#print(unmatched_saudi)
```

View Greek-associated areas
```{r}
# Filter for Greece unmatched entries
unmatched_greece <- unmatched %>%
  filter(str_detect(location, regex("Crete|Region of Central Macedonia, Greece", ignore_case = TRUE)))

# Preview the unmatched Greece entries
#print(unmatched_greece)
```
Sudan
```{r}
# Filter for Sudan unmatched entries
unmatched_sudan <- unmatched %>%
  filter(str_detect(location, regex("Gedaref State, Sudan|Gezira state, Sudan", ignore_case = TRUE)))

# Preview the unmatched Sudan entries
#print(unmatched_sudan)
```

Nigeria
```{r}
# Filter for Nigeria unmatched entries
unmatched_nigeria <- unmatched %>%
  filter(str_detect(location, regex("Benin City, Nigeria|Benin, Nigeria", ignore_case = TRUE)))

# Preview the unmatched Nigeria entries
#print(unmatched_nigeria)
```

Sao Tome and Principe
```{r}
# Filter for STP unmatched entries
unmatched_stp <- unmatched %>%
  filter(str_detect(location, regex("Principe|Principe Island, West Africa", ignore_case = TRUE)))

# Preview the unmatched STP entries
#print(unmatched_stp)
```

UPDATED FINAL CODE
```{r}
# Process abstracts with new matched columns
abstracts_data_new <- abstracts_data_clean %>%
  filter(!is.na(location) & location != "") %>%
  separate_rows(location, sep = ";\\s*") %>%  # Split multiple locations into rows
  mutate(
    Iso3 = case_when(
      location %in% unmatched_us$location ~ "USA",  # US column
      location %in% unmatched_uk$location ~ "GBR",  
      location %in% unmatched_usstates$location ~ "USA", # US states and other places  
      location %in% unmatched_china$location ~ "CHN", 
      location %in% unmatched_canada$location ~ "CAN", 
      location %in% unmatched_japan$location ~ "JPN",
      location %in% unmatched_australia$location  ~ "AUS",
      location %in% unmatched_greece$location ~ "GRC",
      location %in% unmatched_saudi$location ~ "SAU",
      location %in% unmatched_micronesia$location ~ "FSM",
      location %in% unmatched_switzerland$location ~ "CHE",
      location %in% unmatched_tanzania$location ~ "TZA",
      location %in% unmatched_korea$location ~ "KOR",
      location %in% unmatched_france$location ~ "FRA",
      location %in% unmatched_spain$location ~ "ESP",
      location %in% unmatched_indonesia$location ~ "IDN",
      location %in% unmatched_thai$location ~ "THA",
      location %in% unmatched_italy$location ~ "ITA",
      location %in% unmatched_india$location ~ "IND",
      location %in% unmatched_yemen$location ~ "YEM",
      location %in% unmatched_brazil$location ~ "BRA",
      location %in% unmatched_sudan$location ~ "SDN",
      location %in% unmatched_nigeria$location ~ "NGA",
      location %in% unmatched_stp$location ~ "STP",
      location == "Chang Gung Memorial Hospital" ~ "TWN", 
      
      location == "Boston, Massachusetts" ~ "USA", 
      location == "London, UK" ~ "GBR", 
      location == "Sana'a, Yemen" ~ "YEM", 
      location == "England" ~ "GBR", 
      location == "Wales" ~ "GBR",
      location == "England and Wales" ~ "GBR",
      location == "Birmingham, UK" ~ "GBR",
      location == "England, UK" ~ "GBR",
      location == "Northumberland, UK" ~ "GBR",
      
      location == "Glasgow, Scotland" ~ "GBR",
      location == "Central Lancashire, UK" ~ "GBR",
      location == "Washington, D.C." ~ "USA",
      location == "Michigan" ~ "USA",
      
      location == "Virginia, USA" ~ "USA",
      location == "NHS England" ~ "GBR",
      location == "Cabell County, West Virginia, USA" ~ "USA",
      location == "Huntington, West Virginia, USA" ~ "USA",
      
      location == "Congo RDC" ~ "COD",
      location == "Michigan, US" ~ "USA",
      location == "Children's Clinical University Hospital" ~ "LVA",
      location == "Virginia Commonwealth University, USA" ~ "USA",
      
      
      location == "Rakai" ~ "UGA",
      location == "Montgomery County, Virginia" ~ "USA",
      location == "South of England" ~ "GBR",
      location == "Montgomery County, Virginia, USA" ~ "USA",
      
      
      location == "Utah" ~ "USA",
      location == "Arizona" ~ "USA",
      location == "Navajo Nation" ~ "USA",
      location == "New York state, America" ~ "USA",
      location == "Gezira state, Sudan" ~ "SDN",
      location == "Chang Gung Memorial Hospital" ~ "TWN",
      location == "North Wales, UK" ~ "GBR",
      location == "Masaika" ~ "TZA",
      location == "Dundee, UK" ~ "GBR",
      location == "Plymouth, UK" ~ "GBR",
      location == "Nottingham, UK" ~ "GBR",
      location == "East London, UK" ~ "GBR",
      location == "Bristol, UK" ~ "GBR",
      location == "Wisconsin" ~ "USA",
      location == "Northwest England" ~ "GBR",
      location == "East Anglia, UK" ~ "GBR",
      location == "Leicestershire, England" ~ "GBR",
      location == "Eyam, Derbyshire, England" ~ "GBR",
      location == "Greater Manchester, UK" ~ "GBR",
      location == "Addenbrooke's Hospital, UK" ~ "GBR",
      location == "Southern California" ~ "USA",
      location == "Norway House, Manitoba, Canada" ~ "CAN",
      location == "Pemba Island, Zanzibar" ~ "TZA",
      location == "Stony Brook University Hospital, New York" ~ "USA",
      
      location == "Molise" ~ "ITA",
      location == "Basilicata" ~ "ITA",
      location == "Calabria" ~ "ITA",
      location == "Abruzzo" ~ "ITA",
      location == "Sicily" ~ "ITA",
      location == "Apulia" ~ "ITA",
      location == "Campania" ~ "ITA",
      
      
      location == "Lazio" ~ "ITA",
      location == "Tuscany" ~ "ITA",
      location == "Emilia Romagna" ~ "ITA",
      location == "Veneto" ~ "ITA",
      location == "Piedmont" ~ "ITA",
      location == "Honghu, China" ~ "CHN",
      location == "Lombardia" ~ "ITA",
      
      
      location == "Arizona, U.S." ~ "USA",
      location == "Texas, U.S." ~ "USA",
      location == "Colorado, U.S." ~ "USA",
      location == "Charlottesville, Virginia, USA" ~ "USA",
      location == "New York, U.S." ~ "USA",
      location == "New Hampshire, U.S." ~ "USA",
      location == "Taizhou Enze Medical Center (Group) Enze Hospital, China" ~ "CHN",
      location == "Washington" ~ "USA",
      location == "Dundee, Scotland" ~ "GBP",
      location == "Taizhou Enze Medical Center (Group) Enze Hospital, China" ~ "CHN",
      location == "Czechoslovakia" ~ "NA",
      location == "former Czechoslovakia" ~ "NA",
      location == "Congo RDC" ~ "COD",
      location == "Durban" ~ "ZAF",
      location == "Ferlo" ~ "SEN", 
      location == "Yugoslavia" ~ "SRB",  
      location == "San Juan, PR" ~ "PRI",  
      location == "Omsk, Western Siberia" ~ "RUS",
      location == "Russian Federation" ~ "RUS", 
      location == "Russia" ~ "RUS",
      location == "Nile Delta" ~ "EGY",  
      location == "Southwestern Vietnam" ~ "VNM",  
      location == "Southern Vietnam" ~ "VNM",
      location == "Eire" ~ "IRL", 
      location == "Pomerania" ~ "POL", 
      location == "Dutch administrative areas" ~ "NLD",  
      location == "Bioko Island" ~ "GNQ",  
      location == "Bavaria" ~ "DEU",
      location == "German Ruhr Area" ~ "DEU",
      location == "Khuzestan Province" ~ "IRN",
      location == "Madeira" ~ "PRT",
      location == "Malindi District" ~ "KEN",
      location == "Oslo University Hospital" ~ "NOR",
      location == "Auckland" ~ "NZL",
      location == "Saxony" ~ "DEU",
      location == "Al-Bahrani" ~ "ARE",
      location == "D.Z.)" ~ "DZA",
      location == " D.Z.)" ~ "DZA",
      location == "I.S." ~ "ISL",
      location == "Transkei, Southeast Africa" ~ "ZAF",
      location == "Tecumseh" ~ "CAN",
      location == "MS" ~ "MSR",
      location == "Haseki Research and Training" ~ "TUR",
      location == "Kruger National Park" ~ "ZAF",
      location == "Volta River basin area" ~ "GHA",
      location == "SLN Medical College and Hospital" ~ "IND",
      location == "Oxford region" ~ "GBR",
      location == "Rakai" ~ "UGA",
      location == "Kingwede, East Africa" ~ "KEN",
      location == "Sun Yat-sen University Cancer Center" ~ "CHN",
      location == " Al-Bahrani" ~ "ARE",
      location == "Clinic for Nephrology and Clinical Immunology of the Clinical Centre of Vojvodina" ~ "SRB",
      TRUE ~ countrycode::countrycode(
        sourcevar = location,
        origin = "country.name",
        destination = "iso3c",
        warn = FALSE  # Suppress warnings for unmatched entries
      )
    ),
    Iso3 = ifelse(is.na(Iso3), "unmapped", Iso3)  # Mark unmatched values as "unmapped"
  )

# Identify unmatched values for review
unmatched_new <- abstracts_data_new %>%
  filter(Iso3 == "unmapped") %>%
  distinct(location) %>%
  arrange(location)

# Print unmatched values for verification
#print(unmatched_new)
```

```{r}
# Identify newly matched values
newly_matched <- abstracts_data_new %>%
  filter(location %in% unmatched$location, Iso3 != "unmapped") %>%
  distinct(location, Iso3)

# Preview the newly matched values
#print(newly_matched)
```
Now we have as many matched values as we could. We will now combine matched (from part 2) and newly_matched so that we can map everything onto the plot.

```{r}
# Combine newly_matched with matched
all_matched <- bind_rows(matched, newly_matched)
```

```{r}
#these are all the abstracts which we don't map
abstracts_data_clean_unmatched <- abstracts_data_new %>%
  filter(Iso3 == "unmapped")

write.csv(abstracts_data_new, "data/clean_idm_lit_with_locations.csv",row.names = FALSE)
```


```{r}

# Summarize by Iso3 (count the number of abstracts per country)
all_matched_summary <- abstracts_data_new %>%
  group_by(Iso3) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(all_matched_summary, by = c("iso_a3" = "Iso3"))

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = count)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Abstract Count") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/abstract_locations.jpg")


```
