---
title: "R Markdown (Figure Merging)"
author: "Emma Chapman-Banks"
date: "2025-01-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ***Part 1: Merging Figures 1 and 4***

Load necessary packages
```{r}
library(ggplot2)
library(dplyr)
library(janitor)
```

First, we will read, clean, and save publication data. This is from Paula's original code 
```{r}
# Get Publication data
df <- read.csv("data/490k_141224_1234_dedup.csv") #Or change file path to wherever files are located
df <- clean_names(df)

# Remove all excluded records
df <- df |>
  filter(decision == "Include")

# Calc publication records by year 
df_annual <- df |>
  group_by(year) |>
  summarize(n = n()) |>
  rename(n_pub = n)

# Remove records prior to 1900 or after 2025 
df_annual <- df_annual |>
  filter(year < 2025) |>
  filter(year > 1900)

df_annual_save <- df_annual
```

Now, we will load in data for policy citations. Again, this is Paula's original code 
```{r}

pc <- read.csv("data/overton_results_expanded.csv") 
pc <- clean_names(pc)

# Transform date of policy documents to year get year 
pc$published_on <- as.Date(pc$published_on)
pc$year <- format(as.Date(pc$published_on, format="%Y-%m/%d"),"%Y")

# Calc policy citations by year 
pc_annual <- pc |>
  group_by(year) |>
  summarize(n = n()) |>
  rename(n_cit = n)

# Calc policy citations by org type and subtype
pc_subtype <- pc |>
                group_by(type, subtype) |>
                summarize(n = n()) |>
                rename(n_cit = n)

write.csv(pc_subtype, 
          "plots/ec_aggregation_by_sub_type.csv", 
          row.names = FALSE)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))

# Join the summary data with the world map
# world_with_data <- world %>%
#   mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
#   left_join(pc, by = c("iso_a3" = "country_iso3"))
# 
# world_with_data <- world_with_data %>%
#   dplyr::filter(!is.na(published_on))
# 
# world_with_data$type.y <- factor(world_with_data$type.y, levels = c("government", "think tank", "other"))
# 
# ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
#   geom_bar(stat = "identity") +
#   scale_fill_viridis_d(option = "D", name = "Income Group") + # Use the discrete version of viridis
#   labs(
#     x = "Year", 
#     y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
#   ) + 
#   facet_grid(rows = vars(type.y)) + # Stack facets vertically based on type.y
#   theme_minimal() +
#   theme(
#     axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis text by 45 degrees
#   )
# 
# ggsave("plots/by_income_group_type.png",
#        width = 14)

```

Let's first plot the plots separately to see what we are working with. Still Paula's code. 
```{r}
#Plot IDM in Policy Documents 
ggplot(pc_annual, aes(x = as.numeric(year), y = n_cit)) +
  geom_line() +
  labs(x = "Year", y = "Citations of Infectious Disease Modelling Papers in Policy Documents") +
  theme_bw()

# Check the structure and summary of the data
str(pc_annual)
summary(pc_annual)

# Find rows with missing or invalid values
pc_annual |> filter(is.na(n_cit) | n_cit <= 0)
```

Plot publication data. 
```{r}
#Plot Publication Count 
ggplot(df_annual, aes(x = year, y = n_pub)) +
  geom_line() +
  labs(x = "Year", y = "Publication Count") +
  theme_bw()
```

Let's double check the structure of the dataset for publication count too. 
```{r}
# Check the structure and summary of the data
str(df_annual)
summary(df_annual)
```
It seems that pc_annual starts from 1989 and df_annual starts from 1948. This can create issues when merging so we must align the data ranges to ensure RStudio can read this correctly. 
```{r}
# Merge datasets, keeping all years from df_annual
combined_df <- merge(df_annual, pc_annual, by = "year", all.x = TRUE)

# Replace missing policy citation values (n_cit) with 0 (as pc_annual doesn't begin until 1989)
combined_df$n_cit[is.na(combined_df$n_cit)] <- 0

# Ensure n_pub is numeric
combined_df$n_pub <- as.numeric(combined_df$n_pub)

# Plot with secondary y-axis
ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count")) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations")) +  # Scale factor for secondary axis
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations")
  ) +
  labs(x = "Year", color = "Legend") +
  theme_bw() +
  theme(
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    legend.position = "bottom"
  )
```

# ***Part 2: Merging all three plots together***

Now, we will "superimpose" the publication count/policy citations with the outbreak timeline. 

But first, lets load the outbreak timeline data (Paula's code)

```{r}
# Get Outbreaks timeline data 
e_timeline <- readxl::read_excel("data/Epidemics Timeline.xlsx", 
                                 sheet = 2)
names(e_timeline)[1] <- c("year")
e_timeline <- clean_names(e_timeline)

# Reshape data to long format for ggplot2 
data_long <- tidyr::pivot_longer(e_timeline, 
                                 -year, 
                                 names_to = "pathogen", 
                                 values_to = "presence")
data_long <- data_long |> filter(presence > 0)
data_long <- data_long |> filter(year >= min(df_annual$year))

# Renaming pathogens. This is done to help easier visualisation on the plot 
rename_mapping <- c(
  "zika" = "Zika",
  "swine_flu_h1n1_pandemic" = "H1N1 Swine Flu",
  "severe_acute_respiratory_syndrome_sars_coronavirus" = "SARS",
  "russian_flu_pandemic_h1n1" = "H1N1 Russian Flu",
  "m_pox" = "MPox",
  "hong_kong_flu_pandemic_h3n2" = "H3N2 Hong Kong Flu",
  "asian_flu_pandemic_h2n2" = "H2N2 Asian Flu",
  "mers" = "MERS",
  "covid_19" = "COVID-19",
  "ebola" = "Ebola",
  "uptick_in_polio" = "Polio",
  "hiv_aids_pandemic" = "HIV/AIDS",
  "cholera" = "Cholera"
)

data_long$pathogen <- dplyr::recode(data_long$pathogen, !!!rename_mapping)

# Reorder factor levels by frequency 
data_long$pathogen <- factor(data_long$pathogen, 
                             levels = names(sort(table(data_long$pathogen), 
                                                 decreasing = TRUE)))

# Order dataframe by year
data_long <- data_long[order(data_long$year), ]

# Plot Outbreak Timeline
ggplot(data_long, aes(x = year, y = pathogen, color = pathogen)) +
  geom_line(stat = "identity", size = 2) +
  scale_x_continuous(breaks = seq(min(e_timeline$year), max(e_timeline$year), by = 2)) +
  labs(y = "Outbreak") +
  theme_minimal() +
  theme(legend.position = "none", 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(), 
        axis.title.x = element_blank())
```

We will now add the outbreak plot onto the publication/policy plot.

This is the old plot (no disease y-axis with circle dot points)
```{r}
#Because I had issues with creating the legend. I needed to convert the legend labels into a factor (for pathogens) then precompute the breaks 
data_long$pathogen <- factor(data_long$pathogen)
pathogen_breaks <- c("Publication Count", "Policy Citations", levels(data_long$pathogen))

# Create the plot
ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1) +
  geom_point(
    data = data_long, 
    aes(x = year, y = as.numeric(as.factor(pathogen)) * 300, color = pathogen), 
    size = 3, alpha = 0.5
  ) +
  # Scales
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(
    name = "Legend",
    values = c(
      "Publication Count" = "blue",
      "Policy Citations" = "red",
      "Zika" = "lightpink",
      "H1N1 Swine" = "hotpink",
      "SARS" = "violet",
      "H1N1 Russian" = "mediumpurple1",
      "MPox" = "lightblue2",
      "H3N2 Hong Kong" = "azure",
      "H2N2 Asian" = "cyan4",
      "MERS" = "lightgreen",
      "COVID-19" = "mediumaquamarine",
      "Ebola" = "palegreen4",
      "Polio" = "maroon3",
      "HIV/AIDS" = "wheat1",
      "Cholera" = "lightsalmon"
    ),
    breaks = pathogen_breaks
  ) +
  labs(x = "Year", color = "Legend", y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 8, face = "bold"),
    legend.text = element_text(size = 6),
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    axis.text.y.left = element_text(size = 8),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.25), 
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.25) 
  ) +
  guides(
    color = guide_legend(
      override.aes = list(size = 3),
      nrow = 5,
      byrow = TRUE,
      title.position = "top",
      title.hjust = 0.5
    )
  )

```
This is the updated plot (January 22nd 2025) with rectangular bars instead of circular ones. 
```{r}
#Because I had issues with creating the legend. I needed to convert the legend labels into a factor (for pathogens) then precompute the breaks 
data_long <- data_long %>%
  mutate(outbreak_y = as.numeric(as.factor(pathogen)))  # Create a unique numeric value for each pathogen

ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1) +
  geom_tile(
    data = data_long,
    aes(x = year, y = outbreak_y * 300),
    width = 1, height = 200,
    fill = "grey", alpha = 0.5
  ) +
  # Add outbreak labels with fixed colors (excluded from legend with the inhereit.aes = FALSE)
  geom_text(
    data = data_long,
    aes(x = min(combined_df$year) - 5, y = outbreak_y * 300, label = pathogen),
    hjust = 1, size = 3, inherit.aes = FALSE,
    color = "black"
  ) +
  # Scales and labels 
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(
    values = c(
      "Publication Count" = "#35B779FF",
      "Policy Citations" = "#31688EFF"
    )
  ) +
  labs(x = "Year", color = "Legend", y = NULL) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.25),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.25)
  )
```
This is annother updated plot (January 23rd 2025) with the bars being assigned with a specific colour. 
```{r}
data_long <- data_long %>%
  mutate(outbreak_y = as.numeric(as.factor(pathogen)))  # Create a unique numeric value for each pathogen

ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1) +
  geom_tile(
    data = data_long,
    aes(x = year, y = outbreak_y * 300, fill = pathogen), 
    width = 1, height = 200,
    alpha = 0.5
  ) +
  geom_text(
    data = data_long,
    aes(x = min(combined_df$year) - 5, y = outbreak_y * 300, label = pathogen),
    hjust = 1, size = 3, inherit.aes = FALSE,
    color = "black"
  ) +
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(
    values = c(
      "Publication Count" = "#35B779FF",
      "Policy Citations" = "#31688EFF"
    )
  ) +
  scale_fill_viridis_d(name = "Pathogens", option = "plasma") + # This is where the Viridis color scale is applied
  labs(x = "Year", color = "Legend", fill = "Outbreaks") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.25),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.25)
  )
```

Updated Plot (Jan 24th): adding labels to the bars. 

Issue here: the horizontal bars are made up of multiple individual rectangles. You cannot see it in the graph because it has been adjusted so that each rectangle is the same height and width. Therefore, what I have done is pass a code first the calculates the central y position and average x position of each "rectangle". So you must pass this code before plotting it. 
```{r}
# Making sure data is numeric 
data_long <- data_long %>%
  mutate(outbreak_y = as.numeric(as.factor(pathogen))) 

first_appearance <- data_long %>%
  group_by(pathogen) %>%
  summarise(first_year = min(year)) %>%
  arrange(first_year)

# Apply new order to pathogen factor levels
data_long$pathogen <- factor(data_long$pathogen, levels = first_appearance$pathogen)

data_long <- data_long %>%
  mutate(
    disease_characteristic = case_when(
      pathogen %in% c("COVID-19", "SARS", "MERS", 
                      "H1N1 Swine Flu", "H1N1 Russian Flu", 
                      "H3N2 Hong Kong Flu", "H2N2 Asian Flu") ~ "Respiratory",
      pathogen %in% c("Zika", "MPox", "Ebola") ~ "Vector-Borne/Zoonotic",
      pathogen %in% c("HIV/AIDS", "MPox") ~ "STD",
      pathogen %in% c("Cholera") ~ "Waterborne",
      pathogen %in% c("Zika", "Polio") ~ "Neurological Impact",
      pathogen %in% c("Ebola") ~ "Hemorrhagic",
      TRUE ~ "Other" # Default category if none of the above match
    )
  )

# Calculate the central position for the labels
centered_data <- data_long %>%
  group_by(pathogen) %>%
  summarise(
    center_y = mean(outbreak_y * 300),  # This is the center of the horizontal bar
    center_x = mean(year)              # This is the center x position (average year)
  )

centered_data$center_x <- ifelse(centered_data$pathogen == "Ebola", 2016.4, centered_data$center_x)

ebola_add <- data.frame(pathogen = "Ebola",
                        center_y = 1800,
                        center_x = 2000)
centered_data <- rbind(centered_data, ebola_add)

centered_data$pathogen <- factor(centered_data$pathogen, levels = first_appearance$pathogen)

# Plot with new  labels
ggplot(combined_df, aes(x = year)) +
  geom_line(aes(y = n_pub, color = "Publication Count"), size = 1) +
  geom_line(aes(y = n_cit * 15, color = "Policy Citations"), size = 1) +
  geom_tile(
    data = data_long,
    aes(x = year, y = outbreak_y * 300, fill = disease_characteristic), 
    width = 1, height = 200,
    alpha = 0.5
  ) +
  geom_text(
    data = centered_data,
    aes(
      x = center_x,          
      y = center_y,          
      label = pathogen
    ),
    vjust = 0.5,             
    hjust = 0.5,             
    size = 3,
    inherit.aes = FALSE,
    color = "black"
  ) +
  scale_y_continuous(
    name = "Publication Count",
    sec.axis = sec_axis(~ . / 15, name = "Policy Citations"),
    expand = expansion(mult = c(0.1, 0.1))
  ) +
  scale_x_continuous(expand = expansion(mult = c(0.1, 0.1))) +
  scale_color_manual(
    values = c(
      "Publication Count" = "#35B779FF",
      "Policy Citations" = "#31688EFF"
    )
  ) +
  scale_fill_viridis_d(name = " ", option = "plasma") +
  labs(x = "Year", color = " ", fill = "Outbreaks") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.title.y.right = element_text(color = "black"),
    axis.text.y.right = element_text(color = "black"),
    panel.grid.major = element_line(color = "grey80", linewidth = 0.25),
    panel.grid.minor = element_line(color = "grey90", linewidth = 0.25)
  )

ggsave("plots/timeline.png", width = 10)
```

