---
title: "R Markdown (Policy Doc/IDM Lit Analysis )"
author: "Emma Chapman-Banks"
date: "2025-01-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos = c(CRAN = "https://cran.rstudio.com"))
```

# ***Part 1: Cross-Referencing Policy Documents and Abstracts (Same Country)*** 

Load libraries
```{r}
library("tidyverse")
library("sf") 
library("countrycode") 
library("ggrepel") 
library("gt")
library("dplyr")
library("tidyr")
```

Read the data
```{r}
overton_data <- read.csv("data/overton_results_expanded.csv")
idm_data <- read.csv("data/clean_idm_lit_with_locations.csv")

overton_data <- overton_data[, !grepl("^X", names(overton_data))]
idm_data <- idm_data[, !grepl("^X", names(idm_data))]

```

World Data
```{r}
# Get world data 
world <- ne_countries(scale = "small", returnclass = "sf")
world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$iso_a3_eh,
                       world$iso_a3)

world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$sov_a3,
                       world$iso_a3)

# Fix iso_a3 for France
world <- world %>%
  mutate(iso_a3 = ifelse(name == "France", "FRA", iso_a3))

```

Before creating this R markdown, I ran the analysis by joining the datasets together then cross-referencing the dois and country location. This returned zero results. At first, I was puzzled, but then I realised it was because the countries might be spelt differently or be abbreviated. Therefore, before joining the datasets, I used the Iso3 country codes and standardised the country names across the data sets 

Standardise IDM Data. This code is the same code used in the World Mapping Markdown.

When trying to standarise the country codes for Overton data, I also realised that the new column it creates returns a value of NA if the country code is already written. Thus, I identified which country codes were already listed in the dataset and had to include the manual country mapping and ensure it doesn't pass as NA when doing country code mapping. 

```{r}
# Define manual mappings for unmatched values
manual_country_mapping <- c(
  "IGO" = "IGO",  
  "EU" = "EU"     
)

# Standardise Overton Data and include manual mapping
overton_data_clean <- overton_data %>%
  mutate(
    country_iso3 = if_else(
      country %in% names(manual_country_mapping),
      manual_country_mapping[country],
      countrycode::countrycode(
        sourcevar = country,
        origin = "country.name",
        destination = "iso3c",
        warn = FALSE 
      )
    )
  )

```

Now, cross-reference the two datasets.
```{r}
# Combine the datasets based on DOI and use the ISO3 country codes for comparison
combined_data <- overton_data_clean %>%
  inner_join(idm_data, by = c("source_doi" = "doi"))

combined_data <- combined_data[, !grepl("^X", names(combined_data))]

# Filter where the policy document's country matches the IDM abstract's location
matching_countries <- combined_data %>%
  filter(country_iso3 == Iso3) %>%
  filter(Status != "Unmapped")

# Count the number of matches
number_of_matches <- nrow(matching_countries)

# Print the result
print(number_of_matches)

# Summarize by Iso3 (count the number of abstracts per country)
matching_countries_map <- matching_countries %>%
  group_by(Iso3) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(matching_countries_map, by = c("iso_a3" = "Iso3"))

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = count)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Abstract Count") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_from_their_own_countries.jpg")
```

# ***Part 2: Cross-Referencing Policy Documents and Abstracts (Different Country)*** 

Now, let's answer this question: how many policy documents cite IDM literature from a country different from their own?
```{r}
# Filter where the policy document cites IDM literature from a different country
global_research_citations <- combined_data %>%
  filter(country_iso3 != Iso3)  %>%
  filter(Status != "Unmapped")

# Count the number of matches
num_global_citations <- nrow(global_research_citations)

# Print the result
print(num_global_citations)

# Summarize by Iso3 (count the number of abstracts per country)
global_research_citations_map <- global_research_citations %>%
  group_by(Iso3) %>%
  summarize(count = n(), .groups = "drop") %>%
  arrange(desc(count))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(global_research_citations_map, by = c("iso_a3" = "Iso3"))

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = count)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Abstract Count") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_from_other_countries.jpg")
```

# ***Part 3: Deviance between self-citing & other countries*** 

Now, let's answer this question: what's the difference in citing national work versus citing foreign work?
```{r}
# Filter where the policy document cites IDM literature from a different country
names(global_research_citations_map) <- c("Iso3", "count_foreign_citations")
names(matching_countries_map) <- c("Iso3", "count_national_citations")

combined_data_deviance <- merge(matching_countries_map, global_research_citations_map, by = "Iso3", all = TRUE)

combined_data_deviance$diff <- combined_data_deviance$count_national_citations - combined_data_deviance$count_foreign_citations

combined_data_deviance$ratio <- combined_data_deviance$count_national_citations/ combined_data_deviance$count_foreign_citations

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(combined_data_deviance, by = c("iso_a3" = "Iso3"))

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = diff)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Citations from foreign IDM vs. national IDM literature") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_deviance.jpg")

# Plot the world map
ggplot(world_with_data) +
  geom_sf(aes(fill = ratio)) +
  scale_fill_viridis_c(option = "plasma", na.value = "grey90", name = "Ratio of Citations from Foreign IDM to national IDM literature") +
  theme_minimal() +
  theme(legend.position = "bottom")

ggsave("plots/citing_ratio.jpg")
```

```{r}

# Convert the date column to Date format (optional but recommended)
overton_data_clean$published_on <- as.Date(overton_data_clean$published_on)

# Extract the year and overwrite the column
overton_data_clean$published_on <- format(overton_data_clean$published_on, "%Y")

pc <- overton_data_clean |>
  group_by(published_on, country_iso3) |>
  summarize(n = n()) |>
  rename(n_cit = n)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(pc, by = c("iso_a3" = "country_iso3"))

world_with_data <- world_with_data %>%
  dplyr::filter(!is.na(published_on))

ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = economy)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(option = "D", name = "Income Group") + # Use the discrete version of viridis
  labs(x = "Year", y = "Number of Infectious Disease Modelling \nCitations in Policy Documents") + 
  theme_minimal()

ggsave("plots/by_economy.png",
       width = 15)

ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(option = "D", name = "Income Group") + # Use the discrete version of viridis
  labs(x = "Year", y = "Number of Infectious Disease Modelling \nCitations in Policy Documents") + 
  theme_minimal()


ggsave("plots/by_income_group.png",
       width = 15)


```


```{r}
1 - length(unique(overton_data_clean$country))/195
pc <- overton_data_clean |>
  group_by(published_on, country_iso3, type) |>
  summarize(n = n()) |>
  rename(n_cit = n)

pc <- pc %>%
  dplyr::filter(!is.na(published_on))

pc_igo <- pc %>% 
  filter(type == "igo")

pc_igo$income_grp <- "Not country-specific"

names(pc_igo) <- c("published_on", "iso_a3", "type.y","n_cit","income_grp")

# Join the summary data with the world map
world_with_data <- world %>%
  mutate(iso_a3 = ifelse(iso_a3 == "-99", sov_a3, iso_a3)) %>%
  left_join(pc, by = c("iso_a3" = "country_iso3"))

world_with_data <- world_with_data[,c("published_on", "iso_a3", "type.y","n_cit","income_grp")]
world_with_data <- st_drop_geometry(world_with_data)

world_with_data <- bind_rows(world_with_data, pc_igo)

world_with_data <- world_with_data %>%
  dplyr::filter(!is.na(published_on))

world_with_data$type.y <- factor(world_with_data$type.y, levels = c("government", "think tank", "igo","other"))

ggplot(world_with_data, aes(x = published_on, y = n_cit, fill = income_grp)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis_d(option = "D", name = "Income Group") + # Use the discrete version of viridis
  labs(
    x = "Year", 
    y = "Number of Infectious Disease Modelling \nCitations in Policy Documents"
  ) + 
  facet_grid(rows = vars(type.y)) + # Stack facets vertically based on type.y
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis text by 45 degrees
  )

ggsave("plots/by_income_group_type.png",
       width = 14)


```


# ***Part 3: Abstracts with 1+ countries *** 

Now, let's look at how many papers covered 1 country, 2 countries or 3 countries. 
```{r}
# Step 1: Split the countries into separate rows
idm_data_split <- idm_data_clean %>%
  mutate(countries = strsplit(as.character(Iso3), ";")) %>%
  unnest(Iso3) %>%
  mutate(Iso3 = trimws(Iso3))  # Clean up extra spaces

# Step 2: Count the number of countries per paper (group by paper ID)
country_count <- idm_data_split %>%
  group_by(doi) %>%  # Replace `paper_id` with the unique identifier for each paper
  summarize(num_countries = n_distinct(Iso3))

# Step 3: Count how many papers cover 1, 2, 3, etc., countries
coverage_summary <- country_count %>%
  count(num_countries)

# Step 4: Rename columns for better readability
coverage_summary <- coverage_summary %>%
  rename(
    `Number of Countries Covered` = num_countries,
    `Paper Count` = n
  )
# Step 5: View results 
print(coverage_summary)
```
