---
title: "R Markdown (World Mapping for Abstracts)"
author: "Emma Chapman-Banks"
date: "2025-01-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# ***Part 1: Map Abstract Geolocation*** 
Load necessary packages
```{r}
library("tidyverse") 
library("sf") 
library("rnaturalearth") 
library("countrycode")
library("ggrepel") 
```

Get world data/change map projection (from Paula's code)
```{r}
# Get world data 
world <- ne_countries(scale = "small", returnclass = "sf")
world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$iso_a3_eh,
                       world$iso_a3)

world$iso_a3 <- ifelse(world$iso_a3 == "-99", 
                       world$sov_a3,
                       world$iso_a3)

# Change map projection 
world %>%
  st_transform(crs = "+proj=wintri") %>%
  ggplot() +
  geom_sf() +
  coord_sf(datum = NA) + # no graticule
  theme_minimal()
```

Now, we will look at the data from the abstracts. Country data information is found in the llm_location column. When first originally running the code, the following warning came about: 

"Warning: Some values were not matched unambiguously: England & Wales, England and Wales, England, UK, Europe, Kalutara District, Lausanne, New York metropolitan area, sub-Saharan Africa, Washington State"

To fix this, we need to manually map these locations. And these locations are added into the code chunk below (see lines 52-71). Subsaharan Africa and Europe are excluded because it is quite broad, but obviously, up to Paula and others to determine whether they want to include/exclude. Moreover, certain locations that are already valid ISO3 codes are excluded from being remapped and avoids redundant remapping. 


```{r}
# Load new dataset with geolocations 
abstracts_data <- read.csv("data/Combined_Location_With_Metadata_All.csv", stringsAsFactors = FALSE)

# List of all countries (replace with a full list if needed)
countries <- tolower(c(
  "Afghanistan", "Albania", "Algeria", "Andorra", "Angola", "Antigua and Barbuda",
  "Argentina", "Armenia", "Australia", "Austria", "Azerbaijan", "Bahamas", "Bahrain",
  "Bangladesh", "Barbados", "Belarus", "Belgium", "Belize", "Benin", "Bhutan", 
  "Bolivia", "Bosnia and Herzegovina", "Botswana", "Brazil", "Brunei", "Bulgaria",
  "Burkina Faso", "Burundi", "Cabo Verde", "Cambodia", "Cameroon", "Canada",
  "Central African Republic", "Chad", "Chile", "China", "Colombia", "Comoros", 
  "Congo", "Costa Rica", "Croatia", "Cuba", "Cyprus", "Czech Republic", "Denmark", 
  "Djibouti", "Dominica", "Dominican Republic", "Ecuador", "Egypt", "El Salvador", 
  "Equatorial Guinea", "Eritrea", "Estonia", "Eswatini", "Ethiopia", "Fiji", 
  "Finland", "France", "Gabon", "Gambia", "Georgia", "Germany", "Ghana", "Greece", 
  "Grenada", "Guatemala", "Guinea", "Guinea-Bissau", "Guyana", "Haiti", "Honduras", 
  "Hungary", "Iceland", "India", "Indonesia", "Iran", "Iraq", "Ireland", "Israel", 
  "Italy", "Ivory Coast", "Côte d'Ivoire", "Jamaica", "Japan", "Jordan", 
  "Kazakhstan", "Kenya", "Kiribati", "Korea", "Kosovo", "Kuwait", "Kyrgyzstan", 
  "Laos", "Latvia", "Lebanon", "Lesotho", "Liberia", "Libya", "Liechtenstein", 
  "Lithuania", "Luxembourg", "Madagascar", "Malawi", "Malaysia", "Maldives", 
  "Mali", "Malta", "Marshall Islands", "Mauritania", "Mauritius", "Mexico", 
  "Micronesia", "Moldova", "Monaco", "Mongolia", "Montenegro", "Morocco", 
  "Mozambique", "Myanmar", "Namibia", "Nauru", "Nepal", "Netherlands", "New Zealand", 
  "Nicaragua", "Niger", "Nigeria", "North Macedonia", "Norway", "Oman", "Pakistan", 
  "Palau", "Palestine", "Panama", "Papua New Guinea", "Paraguay", "Peru", 
  "Philippines", "Poland", "Portugal", "Qatar", "Romania", "Russia", "Rwanda", 
  "Saint Kitts and Nevis", "Saint Lucia", "Saint Vincent and the Grenadines", 
  "Samoa", "San Marino", "Sao Tome and Principe", "Saudi Arabia", "Senegal", 
  "Serbia", "Seychelles", "Sierra Leone", "Singapore", "Slovakia", "Slovenia", 
  "Solomon Islands", "Somalia", "South Africa", "Spain", "Sri Lanka", "Sudan", 
  "Suriname", "Sweden", "Switzerland", "Syria", "Tajikistan", "Tanzania", 
  "Thailand", "Timor-Leste", "Togo", "Tonga", "Trinidad and Tobago", "Tunisia", 
  "Turkey", "Turkmenistan", "Tuvalu", "Uganda", "Ukraine", "United Arab Emirates", 
  "United Kingdom", "UK", "Great Britain", "Britain", "England", "Scotland", 
  "Wales", "Northern Ireland", "Uruguay", "Uzbekistan", "Vanuatu", "Vatican City", 
  "Venezuela", "Vietnam", "Yemen", "Zambia", "Zimbabwe", 
  "United States", "United States of America", "USA", "U.S.A.", "US", "U.S.", "Democratic Republic of the Congo", "DRC", "Congo-Kinshasa", "Congo (Democratic Republic)", "DR Congo", "République Démocratique du Congo (French variation)", "Congo, The Democratic Republic of"
))

# Function to extract and clean country names
extract_countries <- function(location) {
  if (is.na(location)) return(NA)
  # Split the location string into parts and normalize
  parts <- str_split(location, ",|;")[[1]] %>% 
    str_trim() %>% 
    tolower()
  # Keep only parts that match country names
  matched_countries <- parts[parts %in% countries]
  if (length(matched_countries) > 0) {
    return(paste(unique(str_to_title(matched_countries)), collapse = "; "))
  } else {
    return(NA)
  }
}

# Apply the function to the location column
abstracts_data <- abstracts_data %>%
  mutate(location = sapply(location, extract_countries))

# Define manual mappings only for ambiguous single locations
manual_mapping <- list(
  "England and Wales" = "GBR",
  "England & Wales" = "GBR",
  "Fayoum, Egypt" = "EGY",
  "New York metropolitan area" = "USA",
  "Hebei Province, China" = "CHN",
  "Brasilia, Brazil" = "BRA",
  "Michigan, U.S.A.; New York City, U.S.A." = "USA",
  "Edmonton, Alberta, Canada" = "CAN",
  "South India" = "IND",
  "sub-Saharan Africa" = "unmapped",  # Exclude broad regions
  "southeastern United States" = "USA",
  "Kalutara District; Sri Lanka" = "LKA",
  "Lausanne" = "CHE",
  "Britain" = "GBR",
  "England, UK" = "GBR",
  "U.S." = "USA",
  "Mexico; Southern region of Mexico" = "MEX",
  "Washington State" = "USA",
  "Europe" = "unmapped",
  "Kalutara District" = "LKA"
)

# Process `llm_location` column
abstracts_data_clean <- abstracts_data %>%
  filter(!is.na(location) & location != "") %>%  
  separate_rows(location, sep = ";\\s*") %>%         
  mutate(
    llm_location_resolved = recode(location, !!!manual_mapping, .default = location),
    Iso3 = ifelse(
      llm_location_resolved %in% c("BRA", "CAN", "CHE", "CHN", "EGY", "GBR", "IND", "LKA"), # Skip remapping for valid ISO3 codes
      llm_location_resolved,
      countrycode::countrycode(
        sourcevar = llm_location_resolved,
        origin = "country.name",
        destination = "iso3c",
        warn = FALSE 
      )
    )
  ) %>%
  mutate(Iso3 = ifelse(is.na(Iso3), "unmapped", Iso3))  # Mark unmatched values as "unmapped"

# Add a column to indicate when a location is mapped or unmapped
abstracts_data_clean <- abstracts_data_clean %>%
  mutate(Status = ifelse(Iso3 == "unmapped", "Unmapped", "Mapped"))

# Identify unmatched values for review
unmatched <- abstracts_data_clean %>%
  filter(Status == "Unmapped") %>%
  distinct(llm_location_resolved) %>%
  arrange(llm_location_resolved)

# Print unmatched values for verification. This is to check and confirm that there hasn't been any countries missed. 
print("Unmatched Values:")
print(unmatched)

# Group by ISO3 code and Status, include unmapped entries
abstracts_by_country <- abstracts_data_clean %>%
  group_by(Iso3, Status) %>%
  summarize(n_abstracts = n(), .groups = "drop") %>%
  mutate(n_abstracts_log = log(n_abstracts + 1))  # Add 1 to avoid log(0) issues

```

```{r}
# Merge with world map
world_with_abstracts <- world %>%
  left_join(abstracts_by_country, by = c("iso_a3" = "Iso3"))
```

```{r}
# Plot the world map
ggplot() +
  geom_sf(data = world, fill = "grey90", color = "grey") +
  geom_sf(data = world_with_abstracts, aes(fill = n_abstracts_log), color = NA) +
  scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = "grey90", 
                      name = "Log of Abstracts") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    axis.text.x = element_blank(),
    legend.position = "bottom",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 8)
  ) +
  labs(
    title = "Geographic Distribution of Infectious Disease Modelling Literature",
    x = NULL, y = NULL
  )
```
